Overall

### 2020-08-04 main_figures/anova.Rmd
added new anova figure

---
title: "Supplementary Figures relative to feature selection"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}

suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(pipeComp)
})
theme_set(theme_cowplot(font_size = 11))
source("../misc_functions.R")
```

# Supplementary Figure `r getFigNb(TRUE)`

```{r, fig.width=6, fig.height=5}
ov <- readRDS("../data/overall.rds")

cl <- ov$evaluation$clustering
cl$norm <- relevel(cl$norm, "norm.seurat")
cl$doubletmethod <- relevel(cl$doubletmethod, "none")
cl$filt <- relevel(cl$filt, "none")
cl$sel <- relevel(cl$sel, "sel.vst")
cl$clustmethod <- relevel(cl$clustmethod, "clust.seurat")
cl$clustmethod <- relevel(cl$clustmethod, "clust.seurat")

# checking for interactions:
# mod <- lm(ARI~resolution*dataset+doubletmethod*sel*filt*norm*clustmethod*dims, data=cl)
# co <- as.data.frame(coef(summary(mod)))
# coi <- co[grep(":",row.names(co)),]
# coi <- coi[grep("dataset", row.names(coi), invert=TRUE),]
# colnames(coi) <- c("Estimate","SE","t","p.value")
# coi <- coi[order(coi$p.value),]

@@

cos2 <- lapply(c(ARI="ARI"), trueK=TRUE, FUN=ff)
aov2 <- dplyr::bind_rows(lapply(cos2, FUN=function(x) x$aov), .id="metric")
p1 <- ggplot(aov, aes(reorder(parameter, -varExpl), varExpl, group=parameter, fill=metric)) + 
  geom_col() + xlab("") + ylab("Proportion of variance explained") + coord_flip()
p2 <- ggplot(aov2, aes(reorder(parameter, -varExpl), varExpl, group=parameter, fill=metric)) + 
  geom_col() + xlab("") + ylab("Variance explained") + theme(legend.position="none") + 
  coord_flip() + ggtitle("ARI at true k")

co <- dplyr::bind_rows(lapply(cos, FUN=function(x) x$coef), .id="metric")
co2 <- dplyr::bind_rows(lapply(cos2, FUN=function(x) x$coef), .id="metric")
for(f in c("norm","filt","doubletmethod","dims","sel","clustmethod")){
  co$coef <- gsub(paste0("^",f),"",co$coef)
  co2$coef <- gsub(paste0("^",f),"",co2$coef)
}
p3 <- ggplot(co, aes(coef, Estimate, group=coef, fill=metric)) + geom_col() + xlab("") +
  ylab("Estimated coefficient") + geom_hline(yintercept = 0, linetype="dashed") + 
  coord_flip() + theme(legend.position="none")
p4 <- ggplot(co2, aes(coef, Estimate, group=coef, fill=metric)) + geom_col() + xlab("") +
  ylab("Estimated coefficient") + geom_hline(yintercept = 0, linetype="dashed") + 
  coord_flip() + theme(legend.position="none")

# plot_grid( plot_grid(p1,p2,rel_widths=c(4,3),labels=LETTERS[1:2]),
#            plot_grid(p3,p4,rel_widths=c(4,3),labels=LETTERS[3:4]),
#            nrow=2, rel_heights=c(3,4))
plot_grid(p1 + ylab("Proportion of variance in metric explained by the parameter"),p2,nrow=2,labels=LETTERS[1:2])
```

### Supplementary Figure `r getFigNb()`

**A:** Proportion of the variance in clustering accuracy metrics explained by main parameters. **B:** Proportion of the variance in Adjusted Rand Index (ARI) explained by the parameters, among the analyses giving the true number of clusters. The ANOVA were performed with the model

`metric ~ dataset*resolution+doubletmethod+filt+sel+norm+clustmethod+dims`

In both panels, the proportion is calculated after substracting the variance attributable to differences due to the dataset, the resolution and their interaction. The alternatives included in the analysis (other than the default ones described in the methods) were: 
```
list( doubletmethod=c("none","doublet.scDblFinder"),
      filt=c("none","filt.default","filt.stringent"),
      norm=c("norm.seurat","norm.sctransform","norm.scran"),
      sel=c("sel.vst","sel.deviance","sel.expr"),
      selnb=c(2000,4000),
      dr=c("seurat.pca"),
      dims=c("elbow", 10, "maxLikGlobal"),
      clustmethod=c("clust.seurat","clust.scran.knnAnnoy") )
```

\newpage

# Supplementary Figure `r getFigNb(TRUE)`

```{r}
rnfn <- function(x){
  x <- gsub("norm\\.","",x)
  x <- gsub("^doublet\\.","",x)
  x <- gsub("> none > ","> no filter > ",x)
  gsub("^none > ","",x)
}

H1 <- renameHrows( scrna_evalPlot_silh(ov, agg.by = c("doubletmethod","filt","norm"), row_split = rnfn(norm),
                                       heatmap_legend_param=list(direction="horizontal"),font_factor = 0.8,
                                       row_names_max_width=unit(7, "cm")), rnfn)
H2 <- renameHrows(
     evalHeatmap(ov, step="clustering", what=c("MI","min_pr"), anno_legend=FALSE, title=c("MI","Min precision"), 
                 agg.by=c("doubletmethod","filt","norm"), font_factor = 0.8, 
                 row_split=rnfn(norm),row_names_max_width=unit(7, "cm")) + 
    evalHeatmap(ov, step="clustering", what=c("ARI"), agg.by=c("doubletmethod","filt","norm"), 
                row_split=rnfn(norm),filter=n_clus==true.nbClusts, title="ARI at true k",
                anno_legend=FALSE, font_factor = 0.8, row_names_max_width=unit(7, "cm")),
  rnfn)

H3 <- renameHrows( scrna_evalPlot_silh(ov, agg.by = c("doubletmethod","sel","norm"), row_split = rnfn(norm),
                                       heatmap_legend_param=list(direction="horizontal"),
                                       font_factor = 0.8, 
                                       row_names_max_width=unit(7, "cm")), rnfn)
H4 <- renameHrows(
     evalHeatmap(ov, step="clustering", what=c("MI","min_pr"), anno_legend=FALSE, title=c("MI","Min precision"),
                 agg.by=c("doubletmethod","sel","norm"), row_split=rnfn(norm), 
                 font_factor = 0.8, row_names_max_width=unit(7, "cm")) + 
    evalHeatmap(ov, step="clustering", what=c("ARI"), agg.by=c("doubletmethod","sel","norm"), 
                row_split=rnfn(norm),filter=n_clus==true.nbClusts, title="ARI at true k", 
                anno_legend=FALSE, font_factor = 0.8, row_names_max_width=unit(7, "cm")),
  rnfn)
```

```{r, fig.width=9, fig.height=8.5}
plot_grid( grid.grabExpr(draw(H1, heatmap_legend_side="bottom")), 
           grid.grabExpr(draw(H2)), nrow=2, labels=LETTERS[1:2])
```

### Supplementary Figure `r getFigNb()`

Silhouette widths of the real subpopulations (**A**) and clustering accuracy (**B**) using different combinations of methods.

\newpage

# Supplementary Figure `r getFigNb(TRUE)`

```{r, fig.width=9, fig.height=8.5}
plot_grid( grid.grabExpr(draw(H3, heatmap_legend_side="bottom")), 
           grid.grabExpr(draw(H4)), nrow=2, labels=LETTERS[1:2])
```

### Supplementary Figure `r getFigNb()`

Silhouette widths of the real subpopulations (**A**) and clustering accuracy (**B**) using different combinations of methods.

>>>>

---
title: "Overall"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(pipeComp)
})
theme_set(theme_cowplot(font_size = 11))
source("../misc_functions.R")
```

```{r, fig.width=4, fig.height=4}
ov <- readRDS("../data/overall.rds")

cl <- ov$evaluation$clustering
cl$norm <- relevel(cl$norm, "norm.seurat")
cl$doubletmethod <- relevel(cl$doubletmethod, "none")
cl$filt <- relevel(cl$filt, "none")
cl$sel <- relevel(cl$sel, "sel.vst")
cl$clustmethod <- relevel(cl$clustmethod, "clust.seurat")

@@

cos2 <- lapply(c(ARI="ARI"), trueK=TRUE, FUN=ff)
aov2 <- dplyr::bind_rows(lapply(cos2, FUN=function(x) x$aov), .id="metric")
aov2 <- aov2[aov2$metric=="ARI",]
aov2$metric <- "ARI at true k"
p <- ggplot(rbind(aov,aov2), aes(reorder(parameter, -varExpl), varExpl, fill=metric, colour=metric)) + 
  geom_col(position="dodge") + xlab("") + ylab("Proportion of variance explained") + 
  coord_flip() + theme(legend.position=c(.62,.85))
pdf("anova.pdf", width=4, height=4)
p
dev.off()
p
```
